---
title: "01_data_import"
output: github_document
---
output: github_document knit后，可生成md文件

this file is for doing data import.

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

Import the first dataset.
janitor作用: 一键统一和清理数据框（data frame）的列名

```{r warning=TRUE}
litter_df = 
  read_csv(
    file  = "data/FAS_litters.csv", 
    skip = 0,
    na = c("","NA",".")
    )

litter_df = janitor::clean_names(litter_df)
```

Import the second dataset.
skip = n, 可跳过前n列

```{r}
pups_df = 
  read_csv(
    file  = "data/FAS_pups.csv", skip = 3,
    na = c("","NA",".")
    )
pups_df = janitor::clean_names(pups_df)
```

##look at the data

```{r}
litter_df
pups_df
```

Head or Tail
查看前n行或者最后n行

```{r}
head(litter_df, 20)
tail(pups_df, 5)
```


Skimming?
对数据框进行快速、全面的数据概览

```{r}
skimr::skim(pups_df)
```


## Oh excel ...

```{r}
mlb_df = read_excel("data/mlb11.xlsx")
```

Look at the data

```{r}
mlb_df
```

load lotr data
impory fotr words

```{r}
fotr_df = 
  read_excel("data/LotR_Words.xlsx",
             range = "B3:D6"
             )
fotr_df
```

```{r}
ttt_df = 
  read_excel("data/LotR_Words.xlsx",
             range = "F3:H6"
             )
ttt_df
```

## import SAS

Read in the PULSE dataset.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")

pulse_df = janitor::clean_names(pulse_df)
pulse_df
```

##what's so bad about "read.csv()"

```{r}
pups_df_base = 
  read.csv("data/FAS_pups.csv")
pups_df_base
```

wont show: lazy about variable names: super low for big dataset
